---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

# Budweiser Data Analysis Report

Refer to [@naivelogic github page](https://github.com/naivelogic/MSDS-6306-Case-Study-01-Budweiser) for details on analysis and [data dictionary](https://github.com/naivelogic/MSDS-6306-Case-Study-01-Budweiser#appendix)

![](https://camo.githubusercontent.com/4b828afbe627428baa8407954d1c7ae7ee4cfdb4/68747470733a2f2f696d616765732e61736b6d656e2e636f6d2f31303830783534302f73706f7274732f66616e617469632f6d7973746572792d6275647765697365722d73757065722d626f776c2d61642d313130323830362d54776f42794f6e652e6a7067)

__Import `beers` dataset__
```{r}
beer = read.csv("https://raw.githubusercontent.com/naivelogic/MSDS-6306-Case-Study-01-Budweiser/master/data/Beers.csv")
head(beer)
```
__Import `Breweries` dataset__
```{r}
brew = read.csv("C:/Users/phhale/OneDrive - Southern Methodist University/MSDS 6306 Doing Data Science/lol/MSDS-6306-Case-Study-01-Budweiser/data/Breweries.csv")
head(brew)
# regions
```

> 1.   How many breweries are present in each state?

```{r}
brew_freq = as.data.frame(sort(table(brew$State),decreasing=TRUE))
names(brew_freq) = c('state', 'freq')
sort(table(brew$State), decreasing=TRUE)
```


```{r}
library(ggplot2)
library(scales)
theme_set(theme_classic())

# inspired - http://r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html


# Plot
ggplot(brew_freq, aes(state, freq)) + geom_bar(stat="identity", width = 0.5, fill="tomato2") + 
   ylab("Count of Breweries in State")+
      labs(title="Bar Chart", 
           subtitle="Number of Breweries by State", 
           caption="Source: Frequency of State from 'brewery' dataset") +
      theme(axis.text.x = element_text(angle=65, vjust=0.6))
```


>2.   Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.

```{r}
df = merge(beer, brew, by.x = "Brewery_id", by.y = "Brew_ID")
head(df, 6)
tail(df, 6)
```

> 3.   Report the number of NA's in each column.

```{r}
# Missing values
colSums(is.na(df))
```

We can see that IBU has 1005 NAs and ABV has 62 NAs. It could be possible to do some mean/median imputation for inference in these missing values. however it would be questionable given this size of the data set and how many are missing especially from `IBU`. 

>4.   Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.


```{r}
library(dplyr)
abv_ibu =df %>%
  group_by(State, region) %>%
  summarise_at(vars("ABV", "IBU"), funs(median(., na.rm=TRUE))) 
head(abv_ibu)
```


```{r}
# Plot
ggplot(abv_ibu, aes(State, ABV)) + geom_bar(stat="identity", width = 0.5, fill="tomato2")+
    ylab("Median ABV")+
      labs(title="Appendix 1 - ABV Median by State", 
           subtitle="DC and Kentucky Produced the Highest Median ABV", 
           caption="Source: Median ABV by States from 'beer and brewery' dataset") +
      theme(axis.text.x = element_text(angle=65, vjust=0.6))
```
DC and Kentucky highest ABV


```{r}
ggplot(abv_ibu, aes(State, IBU)) + geom_bar(stat="identity", width = 0.5, fill="tomato2")+
    ylab("Median IBU")+
      labs(title="Appendix 2 - IBU Median by State", 
           subtitle="Maine Produced the Highest Median IBU", 
           caption="Source: Median IBU by States from 'beer and brewery' dataset") +
      theme(axis.text.x = element_text(angle=65, vjust=0.6))
```
We can tell that IBU median across state is a lot more variabile than ABV. with Maine beung the highest meidain IBU producer. 


```{r}
ggplot(abv_ibu, aes(x = State, y = IBU)) +
  geom_bar(
    aes(color = region, fill = region),
    stat = "identity", position = position_stack()
    )
```



> 5.   Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?


```{r}
abv_ibu2 =df %>%
  group_by(State) %>%
  summarise_at(vars("ABV", "IBU"), funs(max(., na.rm=TRUE)))
#View(abv_ibu2)
```

We can tell that for `ABV` that [MAX]Colorado (0.128) other states with high ABV:Kentucky (0.125), Indiana (0.120) and New York (0.100) are hte highest

We can tell that for `IBU` that [MAX] Oregon (138) other states with high IBU: Virgina (135), Maine (130), Ohio (126).


> 6.   Summary statistics for the ABV variable.


```{r}
print(sd(df$ABV, na.rm=TRUE))
summary(df$ABV)
```
above are the summary statistics. we can see that there are 62 NAs and the standard deviation is 0.0135


> dat%>%
group_by(custid)%>% 
summarise(Mean=mean(value), Max=max(value), Min=min(value), Median=median(value), Std=sd(value))

```{r}

model = aov(ABV ~ IBU,data=df)
summary(model) 


```
Utilizing the Anova variance F-test, for the `ABV ~ IBU` simple linear regression model, above provides sufficient evidence per the p-value <2e-16 indicating there is a relationship between `IBU` and `ABV`
```{r}
plot(model)

```
above are the necessary plots for validating the assumption on this model. Normality (qq plot), equal variance (residual fits) and per description of data design independence (no appearance of violation) - we can conclude the simple linear regression model can be used and there are no violations in the assumption. 

### Person Crrelation 
```{r}
#pearson corr
cor.test(df$ABV, df$IBU, method = 'pearson')
```

Based on the Pearson R statistic for ABV and IBU are highly correlated.


```{r}
ggplot(df, aes(x = IBU, y = ABV)) + 
  geom_point() +
  stat_smooth(method = "aov", col = "red")+
      labs(title="Fitted Regression Plot", 
           subtitle="Assessment of relationship between ABV and IBU", 
           caption="Source: IBU and ABV values are from 'beer and brewery' dataset") 
```
In the above fitted line plot plot we are able to confirm the linear relationship between ABV and IBU as IBU increases so does ABV. This supported the previous correaltion output. 

below is a nicer plot used in the report. same result as above
```{r}
library(ggplot2)
theme_set(theme_minimal())  # pre-set the bw theme.
# inspiration - http://r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html

# Scatterplot
ggplot(df, aes(x = IBU, y = ABV)) + 
  geom_point(aes(col=region)) + 
  geom_smooth(method="aov", col = "red", se=F) + 
  xlim(c(0, 140)) + 
  ylim(c(0.025, 0.13)) + 
  labs(title="Fitted Regression Plot", 
           subtitle="Assessment of relationship between ABV and IBU",
       caption="Source: IBU and ABV values are from 'beer and brewery' dataset") 
```

## Conclusion
American related syle beers are high produced, with further research, we can discover the profitability of each style 
For production consideration, we evaluated that we move east to west the median average of alcohol volume and bitterness increases
Finally, there is substantial evidence to suggest positoive relationship between alcohol volume and bitterness. 
