---
title: "Beer EDA"
author: "Christian Nava"
date: "6/18/2019"
output: 
  rmdformats::readthedown:
    highlight: kate
---

```{r, echo=FALSE}
library(knitr)
library(rmdformats)
library(tidyverse)
library(data.table)

opts_chunk$set(echo = TRUE,
               cache = TRUE,
               prompt = FALSE,
               tidy = TRUE,
               comment = NA,
               message = FALSE,
               warning = FALSE)
opts_knit$set(width = 1000)
```

## Exploratory Data Analysis:

```{r}

# Load beer data from url without header and separate into columns
df_beers <- read.csv("https://raw.githubusercontent.com/BivinSadler/MSDS-6306-Doing-Data-Science/master/Unit%207/Beers.csv", 
                     fill = TRUE,
                     header = TRUE,
                     sep = ",",
                     stringsAsFactors = FALSE)

# Load beer data from url without header and separate into columns
df_breweries <- read.csv("https://raw.githubusercontent.com/BivinSadler/MSDS-6306-Doing-Data-Science/master/Unit%207/Breweries.csv", 
                         fill = TRUE,
                         header = TRUE,
                         sep = ",",
                         stringsAsFactors = FALSE)
```



```{r}
# Merge beer data with breweries data
df_merged <- merge(df_beers, df_breweries, 
                   by.x = "Brewery_id", 
                   by.y = "Brew_ID")
```


```{r}
# Make columns more human readable
df_merged <- df_merged %>% 
  rename(
    Beer_name = Name.x,
    Brewery_name = Name.y)
head(df_merged, 10)
```


## Check for Missing Data

```{r}
a = colnames(df_merged)
b = colSums(is.na(df_merged))  %>% as.data.table

missing_value_table = cbind(a, b)

colnames(missing_value_table) = c("Variables","Missing_values")

missing_value_table = missing_value_table  %>% filter(Missing_values>0)  %>% 
                        mutate("% of Total Values" = round(100 * (Missing_values / nrow(df_merged)), 1))  %>% 
                        arrange(desc(Missing_values))

head(missing_value_table, 10)
```

## Summary Statistics for ABV
```{r}
summary(df_merged$ABV)
```

## Distribution of ABV
```{r}
ggplot(df_merged, aes(x=ABV)) + 
  geom_histogram(aes(y=..density..), colour = "black", fill = "#f1c40e") +
  geom_density(alpha=.2) +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position="none") +
  labs(title = "Distribution of ABV", 
       x = "ABV (%)", 
       y = "Density")
```

## Summary Statistics for IBU
```{r}
summary(df_merged$IBU)
```

## Distribution of IBU
```{r}
ggplot(df_merged, aes(x=IBU)) + 
  geom_histogram(aes(y=..density..), colour = "black", fill = "#39bc9b") +
  geom_density(alpha=.2) +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position="none") +
  labs(title = "Distribution of IBU", 
       x = "International Bitterness Units", 
       y = "Density")
```

## Number of beers by state
```{r}
df_beers_by_state <- df_merged %>% count(State)

ggplot(df_beers_by_state, aes(x=reorder(State, n), 
                 y=n, 
                 fill=State)) + 
  geom_bar(stat="identity") + 
  theme(plot.title = element_text(hjust = 0.5),
        legend.position="none") +
  labs(title = "Number of Beers in Analysis by State", 
       x = "State", 
       y = "Number of Breweries") +
  coord_flip()
```


## Median ABV and IBU by State
```{r}
df_state_medians <- df_merged %>% 
  group_by(State) %>% 
  summarise_at(vars(c("ABV", "IBU")), list(~median(., na.rm = TRUE)))
df_state_medians <- df_state_medians %>% 
  rename(
    median_ABV.state = ABV,
    median_IBU.state = IBU)
df_state_medians
```


## ABV and IBU by bottle/can size
```{r}
df_ounces_count <- df_merged %>% count(Ounces)
df_ounces_medians <- df_merged %>% 
  group_by(Ounces) %>% 
  summarise_at(vars(c("ABV", "IBU")), list(~median(., na.rm = TRUE)))
df_ounces <- merge(df_ounces_count, df_ounces_medians, 
                   by.x = "Ounces",
                   by.y = "Ounces")
df_ounces <- df_ounces %>% 
  rename(
    median_ABV.ounces = ABV,
    median_IBU.ounces = IBU)
df_ounces
```

## Distribution of ounces per container
```{r}
ggplot(df_merged, aes(x=Ounces)) + 
  geom_histogram(aes(y=..density..), colour = "black", fill = "#39bc9b") +
  geom_density(alpha=.2) +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position="none") +
  labs(title = "Distribution of Ounces", 
       x = "Ounces per container", 
       y = "Density")
```


## Number of beers by brewery in descending order
```{r}
df_beers_by_brewery <- df_merged %>% 
  count(Brewery_name)
df_beers_by_brewery[order(-df_beers_by_brewery$n), ]
head(df_beers_by_brewery)
```

## Median ABV & IBU by brewery
```{r}
df_medians_by_brewery <- df_merged %>% 
  group_by(Brewery_name) %>% 
  summarise_at(vars(c("ABV", "IBU")), list(~median(., na.rm = TRUE))) 
df_medians_by_brewery <- df_medians_by_brewery %>% 
  rename(
    median_ABV.brewery = ABV,
    median_IBU.brewery = IBU)
head(df_medians_by_brewery)
```

## Number of beers by city in descending order
```{r}
df_beers_by_city <- df_merged %>% 
  count(City)
df_beers_by_city[order(-df_beers_by_city$n), ]
head(df_beers_by_city)
```

## Median ABV & IBU by city of production
```{r}
df_medians_by_city <- df_merged %>% 
  group_by(City) %>% 
  summarise_at(vars(c("ABV", "IBU")), list(~median(., na.rm = TRUE))) 
df_medians_by_city <- df_medians_by_city %>% 
  rename(
    median_ABV.city = ABV,
    median_IBU.city = IBU)
head(df_medians_by_city)
```









