---
title: "RUnit 9 assignment - Phillip Hale"
output: html_notebook
---


### Part A - Data Clean Up

> 1. Create column for brewery ID that is common to both datasets similar to what you did in the project. So we can merge!
2. Merge the beer and brewery data into a single 
3. Clean the State Column … get rid of extraneous white space.  

```{r}
library(tidyverse)

beer = read.csv("https://raw.githubusercontent.com/BivinSadler/MSDS-6306-Doing-Data-Science/master/Unit%209/Beers.csv")
brew = read.csv("https://raw.githubusercontent.com/BivinSadler/MSDS-6306-Doing-Data-Science/master/Unit%209/Breweries.csv")
df = merge(beer, brew, by.x="Brewery_id", by.y = "Brew_ID")
df$State = str_trim(df$State)
```

> 4. Create One Dataset that has only Colorado and Texas beers and no IBU NAs … name it “beerCOTX”
5. Order beerCOTX by IBU (ascending) ... this will be important later in graphing


```{r}
beerCOTX = subset(df, (df$State == "CO" | df$State == "TX") & !is.na(IBU))
beerCOTX = beerCOTX[order(beerCOTX$IBU),]
```
### B

> 6. Plot ABV v. IBU for both Colorado and Texas (two separate plots) … use ggplot and facets.  

```{r}
library(ggplot2)
ggplot(beerCOTX, aes(IBU, ABV)) +
  geom_point() +
  facet_grid(. ~ State)

```

### C. Model the data
> fit eat state SLR. to assess the relationship between ABV and IBU
Use the regular plot function in base R (not ggplot) to create a scatter plot with the regression line superimposed on the plot.  Again, this should be done for each state.


```{r}
lm_tx = lm(ABV ~IBU, data = subset(beerCOTX, beerCOTX$State == "TX" & !is.na(IBU)))
plot(ABV ~IBU, data = subset(beerCOTX, beerCOTX$State == "TX" & !is.na(IBU)))
abline(lm_tx)
```



```{r}
lm_co = lm(ABV ~IBU, data = subset(beerCOTX, beerCOTX$State == "CO" & !is.na(IBU)))
plot(ABV ~IBU, data = subset(beerCOTX, beerCOTX$State == "CO" & !is.na(IBU)))
abline(lm_co)
```


>ASsumptions

```{r}
par(mfrow = c(2, 2), oma = c(0, 0, 2, 0))
plot(lm_co, which =1:2)
plot(lm_tx, which =1:2)
```
Note: top row = CO and bottom row = TX
normality - assumption met by QQ plot. observations appear to fit linear line. Some positive curve, but nothing unreasonable, TX has more curvature. 
equal variance - fit/residual plot and scal-location show residuals randomly scattered around zero. assumption met
independence - per information provided, nothing to suggest violation. we note assumption that independece, assumption met. 
linear - reviewing the plot from the previous graph, there does appear to be a linear relationship betwen ABV and IBU. assumption met. 


### D. Gain inference from the model
>9. Make sure and print the parameter estimate table.  Interpret the slope of the regression model.  
> 10.  Provide a confidence interval for each slope (from each state).  Provide a sentence that interprets each slope (for each state) but this time include the confidence interval in your interpretation.  

```{r}
summary(lm_tx)
print(confint(lm_tx))
```
__Intrepration of linear model on TX__
- regression equation: ABV = 0.043 + 0.0004172 IBU
- slope interpreation: average ABV increases 0.0004172 for each one unit incerase in IBU in TX
- 95% CI: ABV increases between 0.0003 and 0.0004 for every one unit increas in IBU in TX
- pvalue < 0.05 therefore evidence of linear relationship between IBU and ABV

```{r}
summary(lm_co)
print(confint(lm_co))
```
__Intrepration of linear model on CO__
- regression equation: ABV = 0.047 + 0.0003676 IBU
- slope interpreation: average ABV increases 0.0003676 for each one unit incerase in IBU in CO
- 95% CI: ABV increases between 0.0003 and 0.0004 for every one unit increas in IBU in CO
- pvalue < 0.05 therefore evidence of linear relationship between IBU and ABV

### E. Compare two competing models: External Cross Validation

> 11.  Using the beerCOTX dataframe, add a column to the data that is the square of the IBU column.  Call it IBU2.  Print the head of the dataframe with the new column.  


```{r}
beerCOTX$IBU2 = beerCOTX$IBU^2
head(beerCOTX)
```

> 12. For each state, create a training and test set from the data (60%/40% split respectively).  Print a summary of each new data frame… there should be four: TrainingCO, TestCO, TrainingTX, TestTX.  

```{r}
# https://stackoverflow.com/questions/17200114/how-to-split-data-into-training-testing-sets-using-sample-function

## 60% of the sample size
TXdata = subset(beerCOTX, beerCOTX$State == "CO" & !is.na(IBU))
#COdata = subset(beerCOTX, beerCOTX$State == "TX" & !is.na(IBU))

smp_size <- floor(0.60 * nrow(TXdata))
train_ind <- sample(seq_len(nrow(TXdata)), size = smp_size)

TrainingCO <- TXdata[train_ind, ]
TestCO <- TXdata[-train_ind, ]
print(describe(TrainingCO))
print(describe(TestCO))
```


```{r}

```



```{r}

```

```{r}

```


```{r}

```



```{r}

```

```{r}

```


```{r}

```



```{r}

```

```{r}

```


```{r}

```



```{r}

```

```{r}

```